---
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
# Methods Centered on Determining Predictive Factors {#explanatory}

```{r, echo=FALSE}
pacman::p_load_gh("strengejacke/sjPlot")
pacman::p_load(tidyverse)
pacman::p_load(sjmisc)
pacman::p_load(magrittr)
pacman::p_load(data.table)
pacman::p_load_gh("wilkelab/cowplot")
pacman::p_load(mice)
pacman::p_load(FFTrees)
pacman::p_load(Hmisc)
pacman::p_load(lme4)
pacman::p_load(lmerTest)

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```


```{r}
load("bin/data/labelled_students_Dawson.Rdata")
load("bin/data/course_records_Dawson.Rdata")
```


**Under Construction**

This chapter will be focused on methods which have a sound probabilistic framework, and allow for inference into the statistical importance of predictive factors.

## Details of the data used


```{r,echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# this combines both tables into 1
DT <- as.data.table(dplyr::inner_join(students_last_session, courses, by = "student_number", suffix = c(".last_session", ".courses")))
setkey(DT, student_number, ansession.courses)
str(DT)
# DT[, plot(ansession.last_session, ansession.courses)]
DT[, table(CoteR)]
DT[, table(term.courses)]

DT[student_number=="00001B70E66F950D0F06F7F6F6F63411FF966DE0274EE37469A108A0AAD817C0"][, table(term.courses)]

DT[, Gender := factor(Sexe.courses, levels = c("F", "M"), labels = c("Female", "Male"))]
DT[, table(LangueMaternelle.courses)]
DT[, Language := factor(LangueMaternelle.courses, levels = c("AN", "FR", "AU"), labels = c("anglo", "franco", "other"))]
DT[, table(Language)]
DT[, 'Birth Place' := ifelse(birth_place.courses=="Quebec", "Quebec", "Other")]
DT[, 'Birth Place' := factor(`Birth Place`, levels = c("Other", "Quebec"))]
DT[, table(`Birth Place`)]
setnames(DT, old = "Birth Place", new="Birth_Place")

DT[, table(result, useNA = "always")]
DT[, result := ifelse((result=="" | result=="Department" | result=="Intensive" | result=="Not applicable"), NA, result)]
DT[, table(result, useNA = "always")]
DT[, result := factor(result, levels = c("Passing", "At risk", "Failing"))]
DT[, table(result, useNA = "always")]

str(DT)
DT[, table(prog, useNA = "always")]
DT[, prog := ifelse(prog == "NA", NA, prog)]
DT[, table(prog, useNA = "always")]
DT[, prog := factor(prog)]
DT[, table(prog, useNA = "always")]

DT[, table(program.courses, useNA = "always")]
DT[, table(program.last_session, useNA = "always")]
str(DT)

DT[, quit_next_term := factor(quit.next.term, levels = 0:1, labels = c("No", "Yes"))]

DT[,CoteR := factor(CoteR)]
DT[, table(CoteR, useNA = "always")]
# DT[, hist(Note)]
DT[is.na(Note)]

students_last_session$status %>% table

colnames(DT)
table(DT$course)
DT[, table(`current-out`)]
DT[, table(result)]
DT[, table(status)]
DT$course.dept %>% unique() %>% length()
DT$section %>% unique() %>% length()
DT_describe <- DT[,list(student_number, quit.next.term, quit_next_term, result, Note, MoyenneGroupeEvaluation, section,
                     CoteR, course.dept, prog, `current-out`, status, Gender, Language, Birth_Place, term.courses)]

```



```{r, echo=FALSE}
d <- Hmisc::describe(quit_next_term ~ . , data = DT_describe)
plot(d, what = c("quit_next_term", "result", "Note", "MoyenneGroupeEvaluation",
                     "CoteR", "prog", "current-out", "status", "Gender", "Language", "Birth_Place"))
```

```{r, results='asis', echo=FALSE}
html(d, what = c("quit_next_term", "result", "Note", "MoyenneGroupeEvaluation",
                     "CoteR", "prog", "current-out", "status", "Gender", "Language", "Birth_Place"))
```



## Complete Case Analysis


```{r, eval=TRUE}
DT_use_in_model <- DT[,list(student_number, quit.next.term, Note, quit_next_term,
                            result, prog, Gender, Language, Birth_Place, term.courses)]

cc_ind <- complete.cases(DT_use_in_model)
DT_cc <- DT_use_in_model[cc_ind]



# collapsed by student and term, we took the mean note and mode MSE
DT_collapse <- unique(DT_cc[, list(student_number, quit.next.term, quit_next_term, prog, Gender, Language, Birth_Place, term.courses, mean_note = mean(Note), mode_mse = Mode(result)), by = list(term.courses,student_number)], by = c("student_number","term.courses"))



# DT_cc[student_number=="00001B70E66F950D0F06F7F6F6F63411FF966DE0274EE37469A108A0AAD817C0", list(student_number, quit.next.term, quit_next_term, prog, Gender, Language, Birth_Place, term.courses, Note, mean_note = mean(Note), result, mode_mse = Mode(result)), by = list(term.courses,student_number)]
# 
# DT_collapse[student_number=="00001B70E66F950D0F06F7F6F6F63411FF966DE0274EE37469A108A0AAD817C0"]
# 
# DT_collapse_t1 <- unique(DT_cc[term.courses==1, list(student_number, quit.next.term, quit_next_term, prog, Gender, Language, Birth_Place, term.courses, mean_note = mean(Note), mode_mse = Mode(result)), by = student_number], by = "student_number")


# There were a total of `r nrow(DT_use_in_model)` observations from `r length(unique(DT_use_in_model$student_number))` students. Among these, there were `r sum(cc_ind)` observation with complete data (no missingness) for the variables `quit.next.term`, `Note`, `prog`, `Gender`, `Language` and `Birth_Place`, from `r length(unique(DT_cc$student_number))` students.


```



### Logistic Regression

### Binomial Family

```{r, echo=TRUE}
fits <- lapply(1:5, function(i) glm(quit_next_term ~ Gender + Language + Birth_Place + mean_note + mode_mse, data = DT_collapse[term.courses==i], 
           family = "binomial"))
```

```{r}
# for(i in 1:5) sjPlot::sjp.glm(fits[[i]])
sjPlot::sjp.glm(fits[[1]])
sjPlot::sjp.glm(fits[[2]])
sjPlot::sjp.glm(fits[[3]])
sjPlot::sjp.glm(fits[[4]])
sjPlot::sjp.glm(fits[[5]])
```


```{r, eval=FALSE, echo=FALSE}
knitr::knit_exit()
```


```{r, results='asis', eval=FALSE, echo=FALSE}
sjPlot::sjt.glm(fits[[1]])
```



## Mixed Effects Models

Do day 1 predictions. only look at term 1 records. 

```{r, echo=TRUE, eval=FALSE}
# result = mid semester evalutions
# quit_next_term ~ Gender + Language + Birth_Place + Note + result + (1|student_number)
fit_glmer <- lme4::glmer(quit_next_term ~ Gender + Language + Birth_Place + Note + result + (1|student_number), 
                         data = DT_cc[term.courses == 1],
                         family = binomial)
summary(fit_glmer)
# saveRDS(fit_glmer, file = "models/glmer.rds")
fit_glmer <- readRDS(file = "models/glmer.rds")
```

```{r, echo=FALSE, eval=FALSE}
sjPlot::sjp.glmer(fit_glmer, type = "fe")
```

```{r, results='asis', echo=FALSE, eval=FALSE}
sjPlot::sjt.glmer(fit_glmer)
```








```{r, eval=FALSE, echo=FALSE}
DT[unique(student_number), sort(table(birth_place.courses), decreasing = T), mult = "last"]

# this shows that the last entry for each subject corresponds to the session in students_last_session table
DT[unique(student_number), plot(ansession.last_session, ansession.courses), mult = "last"]
abline(a = 0, b = 1, col = "red")

DT[unique(student_number), plot(ansession.last_session, ansession.courses), mult = "first"]

# distribution of number of rows for each student
DT[, .N, by = "student_number"] %>% select(N) %>% unlist("N") %>% hist
str(DT)
DT[, sjp.xtab(quit.next.term, Sexe.last_session, margin = "row"), mult = "last"]
fit <- glm(quit.next.term ~ Gender + Language + `Birth Place`, data = DT, family = "binomial")
sjPlot::sjp.glm(fit)
summary(fit)
library(lme4)
library(lmerTest)
fit2 <- lme4::glmer(quit.next.term ~ Gender + Language + `Birth Place` + (1|student_number), data = DT, family = binomial)
summary(fit2)
plot(fit2)
ranef(fit2)
sjPlot::sjt.lmer(fit2)
sjPlot::sjp.xtab(DT$student_number, students_last_session$LangueMaternelle, margin = "row")
```






```{r, eval=FALSE, echo=FALSE}
DT_use_in_model$student_number
sjPlot::sjp.xtab(students_last_session$status, students_last_session$Sexe)
sjPlot::sjp.xtab(students_last_session$status, students_last_session$LangueMaternelle, margin = "col")
sjPlot::sjp.xtab(students_last_session$status, students_last_session$LangueMaternelle, margin = "cell")
sjPlot::sjp.xtab(students_last_session$status, students_last_session$LangueMaternelle, margin = "row")

sjPlot::sjp.xtab(students_last_session$status, students_last_session$LangueMaternelle, margin = "cell")
sjPlot::sjp.xtab(students_last_session$status, students_last_session$LangueMaternelle, margin = "row")


courses %>% str

courses[student_number == "00001B70E66F950D0F06F7F6F6F63411FF966DE0274EE37469A108A0AAD817C0"]

students_last_session[student_number == "00001B70E66F950D0F06F7F6F6F63411FF966DE0274EE37469A108A0AAD817C0"]

students_last_session$student_number %>% unique() %>% length()

courses$student_number %>% unique() %>% length()

all(students_last_session$student_number %in% courses$student_number)

unique(DT, by = "student_number")

key(DT)
```

