# Continuing Education at Dawson

```{r conted-local-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,message = FALSE)
knitr::read_chunk("bin/ContEdDawson.R")
```

```{r conted-setup}
```

The department of Continuing at Education offers evening courses across many disciplines throughout the year. The purpose of this report is to give a broad overview of the demographics and sector-level metrics for students taking courses within Cont Ed.

## Services

Herein, we look at the services offered by the Division of Continuing Education.

### Continuing Education's Growth over Time

How have each of the departments increased their ContEd offerings over time? 

```{r dept-level-sizes-over-time}
```

### Department Level Seat Distributions

Which parts of Continuing Education are the most important in 2016? 

```{r dept-level-sizes-2016}
```

### Course Level distributions

If we look at the three most important departments (Math,Humanities and English) in 2016, how are the seats distrinuted across courses?

```{r big-dept-seat-dist-by-course}
```

### Specialized spaces

If we look at the departments that require specialized spaces (i.e. labs), in 2016, how are the seats distributed by course? 
 - Has this evolved over time?
 - What are the seasonal variations?

#### Winter
```{r lab-dept-seat-dist-by-course-winter}
```

#### Fall
```{r lab-dept-seat-dist-by-course-fall}
```

#### Summer
```{r lab-dept-seat-dist-by-course-summer}
```



## Students

### Demographics
Who are the students using continuing education services?



### Success Rates
